### 2015-07-20 text/ms.Rmd
edits from JCW - a few minor typos, and lots of comments and suggestions added at the top.

Ecological resource managers are often faced with reconciling multiple potentially conflicting estimates of population status and trajectory. One solution is to average these predictions in an ensemble, but this approach ignores covariance between models and may not optimally reduce bias and improve accuracy. Superensemble models, commonly used for climate and weather forecasting, may provide a superior solution. Superensembles use the predictions from multiple models as covariates in a new statistical model. Here we evaluate the potential for ensemble and superensemble models ('ensemble methods') to improve estimates of fish population status and trajectory. We fit four data-limited models of population and exploitation dynamics to marine fish stocks and combine their estimates of the mean and slope of biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$) with an ensemble average and three superensembles: a linear model, a random forest, and a boosted regression tree. We selected data-limited models for their applicability across the majority of fish stocks world wide. We build our models on a simulated dataset of 5760 stocks and tested them with cross-validation and a global database of `r ram_stocks_n` stock assessments that have also been assessed with more data-intensive stock assessment models. We find that ensemble methods substantially improve estimates of population status and trajectory. Machine-learning superensembles performed the best for population status: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, rank-order correlation between predicted and true status improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. We found similar improvements when predicting the trajectory of status and when applying the simulation-trained superensembles to catch data for global fish stocks. Ensemble methods have the potential to improve estimates of population status and trajectory across many taxa beyond fish. However, they must be carefully tested, formed from a diverse set of accurate models, and built on a dataset sufficiently representative of the populations to which they are applied.

>>>>

Ecological resource managers are often faced with reconciling multiple potentially conflicting estimates of population status and trajectory. One solution is to average these predictions in an ensemble, but this approach ignores covariance between models and may not optimally reduce bias and improve accuracy. Superensemble models, commonly used for climate and weather forecasting, may provide a superior solution. Superensembles use the predictions from multiple models as covariates in a new statistical model. Here we evaluate the potential for ensemble and superensemble models ('ensemble methods') to improve estimates of fish population status and trajectory. We fit four data-limited models of population and exploitation dynamics to marine fish stocks and combine their estimates of the mean and slope of biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$) with an ensemble average and three superensembles: a linear model, a random forest, and a boosted regression tree. We selected data-limited models for their applicability across the majority of fish stocks world wide. We built our models on a simulated dataset of 5760 stocks and tested them with cross-validation and a global database of `r ram_stocks_n` stock assessments that have also been assessed with more data-intensive stock assessment models. We find that ensemble methods substantially improve estimates of population status and trajectory. Machine-learning superensembles performed the best for population status: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, rank-order correlation between predicted and true status improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. We found similar improvements when predicting the trajectory of status and when applying the simulation-trained superensembles to catch data for global fish stocks. Ensemble methods have the potential to improve estimates of population status and trajectory across many taxa beyond fish. However, they must be carefully tested, formed from a diverse set of accurate models, and built on a dataset sufficiently representative of the populations to which they are applied.
